Copulas and their potential for ecology

### 2019-09-30 Paper.Rmd
edits to Q1 methods and results

\noindent \emph{Type of Submission:} Articles, or Concepts and Synthesis

@@

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data [@Genest2007]. We used the 
implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
in the R programming language. 
We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data 
since many of the copula families we considered include the independent copula. 
If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below for 
information on the families used) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was studied in depth
and recommended by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, was shown to be consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

@@

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described formally elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail associations 
of the families, and provide pictorial descriptions in the Appendices.
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 familes. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 copula, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
We used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case for the 
paremeters.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters.

@@

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of distinct locations. We carried out
pairwise bivariate analyses instead of trying to fit a multivariate copula,
for simplicity and because that approach was
sufficient to answer our research questions; but see the Discussion
for a few words on multivariate copulas. We present
the number of pairs for which a non-normal copula
was the AIC-best copula. We also computed the model-averaged lower- and
upper-tail statistics for each pair of locations, and computed the
mean.

We used a spatial resampling scheme 
(Appendix \ref{SM-spatial_resamp}) to calculate confidence intervals of these means.
The scheme is identical to that proposed by @bjornstadfalck2001 for application in their now 
widely used spline correlogram method; the same scheme was also used by @walteretal2017 
as part of a method for 
performing model selection among matrix regression models.
Code and data for this
project are archived at \url{www.github.com/sghosh89/BIVAN}.

compared AIC weights and tail-dependence of 
minimum-AIC copulas with that of a normal copula. We also counted the number of location pairs 
for which model-averaged lower-tail dependence was 
greater than model-averaged upper-tail dependence 
or vice-versa, and reported the percentage of non-independent 
location pairs showing stronger lower-tail dependence.

@@

Copula structures were often non-normal and 
commonly showed asymmetric tail 
dependence, answering Q1 in the affirmative. 
To make results easier to absorb, we first 
present results for an example bivariate dataset, then we summarize results
for all bivariate datasets, then we present results for an example multivariate dataset, then we 
summarize for all multivariate datasets.

<!--bivariate example, model selection-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
minimum-AIC copula, with AIC weight `r round(max(BivMS_res_Loecke$InfCritRes$AICw),2)`, whereas the 
normal copula had AIC weight essentially 0, answering the 
first part of Q1 for these data. 
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, helping to answer 
the second part 
Q1 for these data. The values also differed substantially 
from each other, helping answer the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

@@

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was not meaningfully large (Table \ref{SM-tab_cc2000fit}).
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- and upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answering the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-sgnificantly. Our 
generally affirmative, empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
<!--***DAN: change the box for Cedar Creek-->
We did the same analysis 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, biomass and Shannon's index were independent; for 1997-1999, results 
were similar to 2000.

@@

Delete the following stuff from sup mat:
Figures S24-27
Tables S7-14, keep Table S15
Tables S16-23, keep S24
S25-32, keep S33
S34-42, keep S42

You will have to go back to the methods and revise those parts
-->

We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-N
for the large majority of location pairs (Table \ref{SM-tab_count_bestcop}), and 
AIC weights for best-fitting copulas (Table \ref{SM-tab_count_AICw}) 
were typically substantially 
larger than normal copula weights (Table \ref{SM-tab_count_AICw_normal}), answering the first part of Q1. Goodness of fit tests 
(Tables \ref{SM-tab_count_p_CvM}, \ref{SM-tab_count_p_KS}) in every case 
failed to reject the hypothesis that the AIC-best copula family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize these data.
Model-averaged lower- (Table \ref{SM-tab_count_LTdep_AICw}) and upper-tail
(Table \ref{SM-tab_count_UTdep_AICw}) dependence statistics were exclusively 
positive and were frequently large, thereby differing 
from what a normal copula would give (i.e., 0), and answering the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from a normal-copula null hypothesis. 
Model-averaged lower- minus upper-tail dependence statistics 
(Table \ref{SM-tab_count_LTmUT}) were positive 
for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This also answers the third part of Q1.
Nonparametric statistics verified that tail dependence was asymmetric for the green 
spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}).

@@

dt <- tibble(Description = c("1. p-value, independence test",
           "2. Minimum-AIC bestfit copula",
           "3. AIC for bestfit copula",
           "4. AIC for normal copula",
           "5. p-value, Cramer-von Mises goodness of fit test",
           "6. p-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates", # this row will be commented out, that's why the numbering starts from 10 in the next line
           "10. Placement of cor$_{FS}$-cor$_{LS}$ in surrogate distribution (see caption)")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
tab_bivar_summary<-tab_bivar_summary[c(1:9,11),]

knitr::kable(tab_bivar_summary, 
             format="latex", escape=F,row.names=F,
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values (rows 5-6) are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. Row 10 shows values as in Table \\ref{tab_soilCN_asym}, upper right table entry. Although the result shown in row 10 for the soil C and N data is non-significant, see Table \\ref{tab_soilCN_asym} for significant results using the $P$ and $D^2$ statistics. The first 3 datasets use $(l_b,u_b)=(0,0.1)$ for the first slice (FS) and $(0.9,1)$ for last slice (LS), whereas $(0, 0.25)$ and $(0.75,1)$ were used for Cedar Creek due to fewer data for that system. Results for Cedar Creek for the year 2000 are shown. See Table \\ref{tab_soilCNfit} for copula family abbreviations.\\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
             )%>%column_spec(2:5,width="1.5cm")%>%
             column_spec(1, width = "8.4cm")%>%kable_styling(latex_options = "basic",position = "center")
```

@@

row.names(tab_multivar_summary)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(
            Description = c("1. Location pairs (excluding self comparisons)",
           "2. Number of non-independent pairs",
           "3. Number of pairs with non-normal copula as best fit",
           #"4. Avg. of normal-AICw minus bestfit-copula AICw across location pairs",
           "4. Average of AIC differences (normal minus bestfit-copula) across location pairs",
           "5. Number of well-fitting location pairs",
           "6. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. lower-tail dependence",
           "7. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. upper-tail dependence",
           "8. Percent pairs showing stronger lower tail dependence",
           "9. 2.5$^{th}$ and 97.5$^{th}$ quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
           )

>>>>

@@

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data [@Genest2007]. We used the 
implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
in the R programming language. 
We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data 
since many of the copula families we considered include the independent copula. 
If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below for 
information on the families used) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was studied in depth
and recommended by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, was shown to be consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using normalized ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

@@

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was solely to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additonally, these copulas have been 
described formally elsewhere [@Joe1997]; so 
we only briefly identify each family, say a few words about tail associations 
of the families, and provide pictorial descriptions in the Appendices.
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 familes. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts that depend on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 for details on all these families. We used the implementations
of these families provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1} and \ref{SM-figped_SC}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters.

@@

For multivariate datasets, we performed the bivariate analysis
for all possible pairwise combinations of distinct locations. We carried out
pairwise bivariate analyses instead of trying to fit a multivariate copula,
for simplicity and because that approach was
sufficient to answer our research questions; but see the Discussion
for a few words on multivariate copulas. 
<!--***DAN: dont forget to put these in!-->
We present
the number of pairs for which a non-normal copula
was the AIC-best copula, and we characterize AIC differences between
the normal and AIC-best copulas across location pairs. 
We also computed the model-averaged lower- and
upper-tail statistics, and differences
between these, for each pair of locations, and we characterize
the distribution of these values across location pairs.

@@

Copula structures were non-normal and 
showed asymmetric tail 
associations for most, but not all datasets we examined, 
answering Q1 in the affirmative. 
To make results easier to absorb, we first 
present results for an example bivariate dataset, then we summarize results
for all bivariate datasets, then we present results for an example multivariate dataset, then we 
summarize for all multivariate datasets.

<!--bivariate example, model selection-->
For the soil C and N data (Data section, Table \ref{tab_data_info}, 
Fig. \ref{fig_biv_multi_raw_cop}A), variables were 
non-independent ($p=$ `r BivMS_res_Loecke$IndepTestRes`, to within the precision available 
from `BiCopIndTest` in the `VineCopula` package); non-independence is also visually apparent.
The Kendall correlation was `r round(BivMS_res_Loecke$TauVal,2)`. We fitted our 16 copulas
to the normalized ranks (Fig. \ref{fig_biv_multi_raw_cop}E) 
and computed AIC and BIC weights and corresponding
lower- and upper-tail statistics for the fitted distributions (Table \ref{tab_soilCNfit}).
`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]` was the 
minimum-AIC copula, with AIC `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)`, whereas the 
normal copula had a much higher AIC, $-6333.16$, 
<!--***DAN: Shyamolina, please make this an autolink. Please make 
sure to round to two digits after the decimal, as with the AIC best value
presented just above.-->
answering the 
first part of Q1 for these data. Recall that AIC differences of $2$
or $3$ are considered meaningful and differences of $8$ or greater
are considered diffinitive, so that data provide essentially no support
for the higher-AIC model in that case [@burnham2003_modelselection].
Model-averaged lower- and upper-tail dependence statistics were `r round(BivMS_res_Loecke$relLTdep_AICw,3)` 
and `r round(BivMS_res_Loecke$relUTdep_AICw,3)`, respectively (AIC weights were used for averaging). 
These values are distinct from what a normal copula would give, namely, 0, helping to answer 
the second part 
Q1 for these data. The values also differed substantially 
from each other, helping answer the third part of Q1. 
The numbers reflect stronger upper- than lower-tail
dependence, and this is also visible in the extreme upper-right
and lower-left corners of the copula plot (Fig. \ref{fig_biv_multi_raw_cop}E).

@@

<!--bivariate summary-->
Table \ref{tab_bivar_summary} summarizes Q1 results 
for the bivariate datasets, with some details in the Appendices. The two 
variables were significantly related for all datasets 
(Table \ref{tab_bivar_summary}, row 1). A non-normal copula 
always emerged as the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, row 2). 
The normal copula was a poor fit compared to the lowest-AIC copula 
(Table \ref{tab_bivar_summary}, rows 3-4), except for Cedar Creek, for which the AIC
difference between these fits was not meaningfully large (Table \ref{SM-tab_cc2000fit}).
Thus we answer the first part of Q1 in the affirmative for 3 of 4 of our bivariate datasets. 
Often, either lower- or upper-tail dependence statistics differed substantially from 0 
(Table \ref{tab_bivar_summary}, rows 7-8), and/or these statistics differed from each other (Table \ref{tab_bivar_summary}, row 9), 
helping to answering the second and third parts of Q1. Though again the model
selection results do not provide 
information on statistical significance of these differences, and are subject to the
caveat that, for some datasets, even the lowest-AIC copula was not an objectively good fit
(Table \ref{tab_bivar_summary}, rows 5 and 6), nonparametric results 
(Table \ref{tab_bivar_summary}, row 10) showed that tail associations
deviated significantly from what would be expected from a normal-copula
null model, and were also significantly asymmetric, except for the Cedar Creek data.
See Figs \ref{SM-fig_bmr_birds_nonparam}-\ref{SM-fig_cc2000_nonparam} for analogues to 
Fig. \ref{fig_soilCN_nonparam} for the bivariate datasets other than the soil C and N
data, and see Table \ref{SM-tab_3biv_npa_summary} for analogues to Table \ref{tab_soilCN_asym}.
The first three datasets had stronger upper- 
than lower-tail dependence, and the last dataset had the reverse, though non-significantly. Our 
generally affirmative (except for Cedar Creek), 
empirically based answer to Q1 for the bivariate datasets is 
represented in Fig. \ref{MasterFigure} as solid outlines around the names of 
the soil C and N and bird and mammal body mass versus metabolic rate
datasets, in the middle row of boxes.
<!--***DAN: change the box for Cedar Creek-->
We did the same analysis 
with Cedar Creek data from other available years, 1996-1999 and 2007. For 1996 and 
2007, biomass and Shannon's index were independent; for 1997-1999, results 
were similar to 2000.

@@

Delete the following stuff from sup mat:
DONE Figures S24-27
DONE Tables S7-14, keep Table S15
DONE Tables S16-23, keep S24
DONE S25-32, keep S33
DONE S34-42, keep S42

DONE You will have to go back to the methods and revise those parts
-->

We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-normal
for the large majority of location pairs (72 of 90 pairs), 
<!--***DAN: Shyamolina, please make these numbers autolinks-->
and the AIC
of the normal copula minus the minimum AIC, averaged across 
location pairs, was $2.732$.
<!--***DAN: Shyamolina, please make this an autolink, so the value is
automatically filled in even if it changes-->
Whereas $2.732$
<!--***DAN: Shyamolna, again, autolink, please-->
would be only a marginally meaningful AIC difference for a single
location pair, it is more meaningful as an average across many
pairs. Relatedly, and illustrating the concept here, 
the chances of getting 72 or more 
non-normal location pairs if the chances were equal of getting
a normal versus a non-normal result (and taking into account that
the location pair (A,B) will necessarily 
produce the same result as the pair
(B,A)) is low, $`r 1-pbinom(72/2-1,45,.5)`$. 
<!--***DAN: Shyamolina, please replace the 72 in the above expression
by an autolink, and make this show
3 significant digits, i.e., it should show 3.29e-5, not 3.287332e-05-->
These results help answer the first part of Q1: non-normal
copula structure appears to be meaningfully common in
these data, even if not universal.
Goodness of fit tests in every case 
failed to reject the hypothesis that the AIC-best copula 
family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize 
these data. Model-averaged lower- and upper-tail
dependence statistics had $2.5$th and $97.5$th quantiles
$(0.123,0.721)$ and $(0.002,0.41)$, respectively,
<!--***DAN: Shyamolina, autolink pls -->
thereby commonly differing 
from what a normal copula would give (i.e., 0), and helping to
answer the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from a normal-copula null hypothesis. We note however that our model selection 
procedures again do not reveal whether tail-dependence 
parameters are significantly different from $0$, and we refer 
the reader to nonparametric results below for information
on significance. 
Model-averaged lower- minus upper-tail dependence statistics 
were positive for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). The chances
of getting `r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
or more positive values, here, under a null hypothesis of equal chances 
for positive and negative values (and again accounting for the fact that location pairs (A,B) and (B,A) will show the same result) was
again low, $`r 1-pbinom(86/2-1,45,.5)`$. 
<!--***DAN: Shyamolina, please autolink this, and use 3 sig figs as 
before-->
Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This also answers the third part of Q1.

<!--***DAN: got to here. Next step, put in the actual nonparam stats results which are currently in the Appendices-->

Nonparametric statistics verified that tail dependence was asymmetric for the green 
spruce aphid abundance data (Appendix \ref{SM-Q1multvarnonparam}, \ref{SM-spatial_resamp}).

@@

dt <- tibble(Description = c("1. p-value, independence test",
           "2. Minimum-AIC copula",
           "3. AIC for best copula",
           "4. AIC for normal copula",
           "5. p-value, Cramer-von Mises goodness of fit test",
           "6. p-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates", # this row will be commented out, that's why the numbering starts from 10 in the next line
           "10. Placement of cor$_{FS}$-cor$_{LS}$ in surrogate distribution (see caption)")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
tab_bivar_summary<-tab_bivar_summary[c(1:9,11),]

knitr::kable(tab_bivar_summary, 
             format="latex", escape=F,row.names=F,
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values (rows 5-6) are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. Row 10 shows values as in Table \\ref{tab_soilCN_asym}, upper right table entry. Although the result shown in row 10 for the soil C and N data is non-significant, see Table \\ref{tab_soilCN_asym} for significant results using the $P$ and $D^2$ statistics. The first 3 datasets use $(l_b,u_b)=(0,0.1)$ for the first slice (FS) and $(0.9,1)$ for the last slice (LS), whereas $(0, 0.25)$ and $(0.75,1)$ were used for Cedar Creek due to fewer data for that system. Results for Cedar Creek for the year 2000 are shown. See Table \\ref{tab_soilCNfit} for copula family abbreviations.\\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
             )%>%column_spec(2:5,width="1.5cm")%>%
             column_spec(1, width = "8.4cm")%>%kable_styling(latex_options = "basic",position = "center")
```

@@

row.names(tab_multivar_summary)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(
            Description = c("1. Location pairs (excluding self comparisons)",
           "2. Number of non-independent pairs",
           "3. Number of pairs with non-normal copula as best fit",
           #"4. Avg. of normal-AICw minus bestfit-copula AICw across location pairs",
           "4. Average of AIC differences (normal AIC minus min AIC) across location pairs",
           "5. Number of well-fitting location pairs",
           "6. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. lower-tail dependence",
           "7. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. upper-tail dependence",
           "8. Percent pairs showing stronger lower tail dependence",
           "9. 2.5$^{th}$ and 97.5$^{th}$ quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
           )
